---
title: "Bibliography"
author: "Lennart Wittkuhn"
date: "Last update: `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    self_contained: true
    toc_float: true
    toc_depth: 3
    number_sections: false
    highlight: pygments
    theme: cosmo
    code_folding: "hide"
    df_print: paged
    fig_caption: true
fig.align: "center"
email: wittkuhn@mpib-berlin.mpg.de
---

```{r, warning=FALSE, message=FALSE, echo=FALSE}
if (!requireNamespace("pacman")) install.packages("pacman")
packages_cran <- c("here", "data.table", "DT", "tidyverse")
pacman::p_load(char = packages_cran)
path_bib <- here::here("public", "bibliography.csv")
```

```{r, echo=FALSE}
bib <- data.table::fread(path_bib) %>%
  # turn URLs into hyperlinks that open in a new tap:
  .[, url := paste0("<a href='", url , "' target='_blank'>URL</a>")] %>%
  # remove curly brackets from the title:
  .[, title := gsub("[{}]", "", title)]
```

# Bibliography

## Overview

Welcome! This website show the entire bibliography maintained by Lennart Wittkuhn.

At the moment, this bibliography contains `r nrow(bib)` references for publications in cognitive and computational neuroscience, psychology, statistics, meta-science and more.

## Usage

The table below lists all references sorted by the first author's last name, publication year, and journal name (in that order).

You can use the `Search` bar on the right to search for certain publications (e.g., by author or journal name).

The `URL` column contains hyperlinks that should bring you to the publisher's website for the corresponding publication.

When you use the `bibliography.bib` file in one of your documents (in LaTeX) and you want to cite one of the references, use the citation key in the `ID` column to cite the relevant publication, e.g., `\cite{Abraham2014FN}`.

If a reference is missing, please [get in touch](#contact).

```{r}
DT::datatable(
  bib, class = "cell-border stripe", rownames = FALSE,
  options = list(scrollX = TRUE), escape = FALSE,
  ) 
```

# Contact

If a reference is missing, please [create a new issue](https://github.com/lnnrtwttkhn/bibliography/issues/new/choose) and use the issue template for missing publications.

If you have any questions about the bibliography, [the repository](https://github.com/lnnrtwttkhn/bibliography/), if you spotted a bug or would like to make a comment, please also [open an issue](https://github.com/lnnrtwttkhn/bibliography/issues/new) first, or otherwise [email Lennart](mailto:wittkuhn@mpib-berlin.mpg.de).

Thanks!

# Packages

This website was built using the following computational environment:

```{r}
sessionInfo()
```


