---
title: "Bibliography"
author: "Lennart Wittkuhn"
date: "Last update: `r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    toc: yes
    self_contained: true
    toc_float: true
    toc_depth: 3
    number_sections: false
    highlight: pygments
    theme: united
    code_folding: "hide"
    df_print: paged
    fig_caption: true
fig.align: "center"
email: wittkuhn@mpib-berlin.mpg.de
---

[![Build Status](https://app.travis-ci.com/lnnrtwttkhn/bibliography.svg?branch=master)](https://app.travis-ci.com/lnnrtwttkhn/bibliography)

```{r, warning=FALSE, message=FALSE, results='hide', echo=FALSE}
if (!requireNamespace("pacman")) install.packages("pacman")
packages_cran <- c(
  "here", "data.table", "DT", "tidyverse", "lemon", "bookdown", "assertr")
pacman::p_load(char = packages_cran)
# save locale:
loc <- Sys.getlocale("LC_TIME")
Sys.setlocale("LC_TIME", "en_US.UTF-8")
```

```{r}
path_bib <- here::here("public", "bibliography.csv")
bib <- data.table::fread(path_bib) %>%
  # turn URLs into hyperlinks that open in a new tap:
  .[, url := paste0("<a href='", url , "' target='_blank'>URL</a>")] %>%
  # remove curly brackets from the title:
  .[, title := gsub("[{}]", "", title)] %>%
  # convert dates when references were added:
  .[, ":="(date_added_converted = as.Date(
  get("date-added"), format = "%Y-%m-%d %H:%M:%S %z"))]
bib_all = bib %>%
  # select columns:
  .[,c("ID", "title", "author", "year", "journal", "url")]
```

# Overview

Welcome! This website showcases the bibliography maintained by Lennart Wittkuhn.

At the moment, this bibliography contains **`r nrow(bib)` references** for publications in neuroscience, psychology, statistics, artificial intelligence, meta-science and more.
The website was last updated on `r format(as.Date(max(bib$date_added_converted), format = "%Y-%m-%d %H:%M:%S %z"), "%d %B, %Y")`.

The source code of this website can be found at <https://github.com/lnnrtwttkhn/bibliography/>.

To understand how the bibliographical data was processed in this document, you can click on the `Code` tabs on the right.
For example, by clicking on the first `Code` tab on the right, you will be able to see the R code that was used to load and process the bibliographical information stored in the `bibliography.bib` file. 

# Usage

The table below lists all references sorted by the first author's last name, publication year, and journal name (in that order).

You can use the `Search` bar on the right to search for certain publications (e.g., by author or journal name).

The `URL` column contains hyperlinks that should bring you to the publisher's website for the corresponding publication.

When you use the `bibliography.bib` file in one of your documents (in LaTeX) and you want to cite one of the references, use the citation key in the `ID` column to cite the relevant publication, e.g., `\cite{Wittkuhn2020B}`.

If a reference is missing, please [get in touch](#contact).

Figure \@ref(fig:bib-complete) shows a table of the entire bibliography.

```{r, bib-complete, fig.cap="Table of all publications in `bibliography.bib`."}
DT::datatable(
  bib_all,
  class = "cell-border stripe", rownames = FALSE, escape = FALSE,
  options = list(
    scrollX = TRUE,
    pageLength = 5,
    lengthMenu = c(5, 10, 15, 20)
    ),
  ) 
```

# Filter

In the sections below, I filter the bibliography e.g., depending on certain keywords.

## Replay

Table \@ref(fig:replay) list all publications in the `bibliography.bib` file that mention the term `replay` in the abstract or the title:

```{r, replay, fig.cap="Table listing studies mentioning 'replay' in the abstract."}
search_terms = c("replay")
bib_replay = bib %>%
  .[, ":="(abstract_lower = tolower(abstract), title_lower = tolower(title))] %>%
  filter_at(
    .vars = vars(abstract_lower, title_lower),
    .vars_predicate = any_vars(
      stringr::str_detect(
        string = .,
        pattern = paste(search_terms, collapse = "|")))) %>%
  setDT(.) %>%
  .[, c("ID", "title", "author", "year", "journal", "url")] %>%
  setorder(author, year)
DT::datatable(
  bib_replay,
  class = "cell-border stripe", rownames = FALSE, escape = FALSE,
  options = list(
    scrollX = TRUE,
    pageLength = 5,
    lengthMenu = c(5, 10, 15, 20)
    ),
  ) 
```

Figure \@ref(fig:replay-year) shows the number of papers published per year, containing replay-related search terms in the abstract.
Note, that the numbers are based only on the publications in the bibliography.
The bibliography likely does not contain *all* publications on `replay` or `reactivation` in the literature.

```{r}
search_terms = c("replay", "reactivation")
bib_replay_time = bib %>%
  transform(abstract = tolower(abstract), title = tolower(title)) %>%
  # search for search terms in title and abstract
  filter_at(
    .vars = vars(abstract, title),
    .vars_predicate = any_vars(
      stringr::str_detect(
        string = .,
        pattern = paste(search_terms, collapse = "|")))) %>%
  setDT(.) %>%
  .[, by = .(year), .(N = .N)] %>%
  setorder(year) %>%
  transform(year = as.Date(ISOdate(year, 1, 1)))
```

```{r, replay-year, fig.cap="Number of papers published per year, containing replay-related search terms in the abstract", fig.fullwidth=TRUE}
title_string = paste("Number of papers published each year\n",
                     "with abstracts containing the following search terms:\n",
                     paste(search_terms, collapse = ", " ))
ggplot(data = bib_replay_time, aes(x = year, y = N)) +
  geom_bar(stat = "identity") +
  ggtitle(title_string) +
  ylab("Number of publications per year") +
  xlab("Publication year") +
  theme(panel.background = element_blank()) +
  theme(panel.grid.major = element_blank()) +
  theme(panel.grid.minor = element_blank()) +
  theme(axis.line = element_line(colour = "black")) +
  coord_capped_cart(bottom = "both", left = "both") +
  scale_x_date(date_breaks = "1 years", date_labels =  "%Y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = N), vjust = -0.5) +
  ylim(c(0, 80))
```

## Replay Literature Reviews

In this section, I filter publications in the `bibliography.bib` file for **literature review papers on replay**.
The matching papers are listed in Figure \@ref(fig:replay-review), sorted by year in descending order (newest to oldest) and author name (alphabetical).

```{r, replay-review, fig.cap="Table of literature reviews on replay."}
bib_replay_review = bib %>%
  filter(stringr::str_detect(string = tags, pattern = "replay")) %>%
  filter(stringr::str_detect(string = tags, pattern = "review")) %>%
  setDT(.) %>%
  .[,c("ID", "title", "author", "year", "journal", "url")] %>%
  setorder(., -year, author)
DT::datatable(
  data = bib_replay_review,
  class = "cell-border stripe", rownames = FALSE, escape = FALSE,
  options = list(
    scrollX = TRUE,
    pageLength = 5,
    lengthMenu = c(5, 10, 15, 20)
  ),
) 
```

## Intracranial recordings in humans

In this section, I filter publications in the `bibliography.bib` file for papers using **intracranial recordings in humans**.
The matching papers are listed in Figure \@ref(fig:human-intracranial).
Many (but not all!) of them are investigating replay-like signals in the medial temporal lobe and therefore establish important correspondence to electrophysiological recordings of replay in rodents.

```{r, human-intracranial, fig.cap="Table listing studies using intracranial recordings in humans."}
bib_intracranial = bib %>%
  filter(stringr::str_detect(string = tags, pattern = "human")) %>%
  filter(stringr::str_detect(string = tags, pattern = "intracranial")) %>%
  setDT(.) %>%
  .[,c("ID", "title", "author", "year", "journal", "url")]
DT::datatable(
  data = bib_intracranial,
  class = "cell-border stripe", rownames = FALSE, escape = FALSE,
  options = list(
    scrollX = TRUE,
    pageLength = 5,
    lengthMenu = c(5, 10, 15, 20)
  ),
) 
```

## Successor representation

```{r}
search_terms = "successor representation"
bib_successor_representation = bib %>%
  .[, ":="(abstract_lower = tolower(abstract), title_lower = tolower(title))] %>%
  # search for search terms in title and abstract
  filter_at(
    .vars = vars(abstract_lower, title_lower),
    .vars_predicate = any_vars(
      stringr::str_detect(
        string = .,
        pattern = paste(search_terms, collapse = "|")))) %>%
  setDT(.) %>%
  .[,c("ID", "title", "author", "year", "journal", "url")]
```

In this section, I filter publications in the `bibliography.bib` file for papers on the **successor representation** in reinforcement learning.
The matching papers are listed in Figure \@ref(fig:successor-representation).
The filter searches for the keyword `successor representation` in the `title`, `abstract` or `tags` of the bibliography entries.
`r nrow(bib_successor_representation)` matching papers were found.

```{r, successor-representation, fig.cap="Table listing studies on the successor-representation in reinforcement learning."}
DT::datatable(
  data = bib_successor_representation,
  class = "cell-border stripe", rownames = FALSE, escape = FALSE,
  options = list(
    scrollX = TRUE,
    pageLength = 5,
    lengthMenu = c(5, 10, 15, 20)
  ),
) 
```

## Cognitive maps

```{r}
search_terms = "cognitive map"
bib_cognitive_map = bib %>%
  .[, ":="(abstract_lower = tolower(abstract), title_lower = tolower(title))] %>%
  # search for search terms in title and abstract
  filter_at(
    .vars = vars(abstract_lower, title_lower, tags),
    .vars_predicate = any_vars(
      stringr::str_detect(
        string = .,
        pattern = paste(search_terms, collapse = "|")))) %>%
  setDT(.) %>%
  .[,c("ID", "title", "author", "year", "journal", "url")]
```

In this section, I filter publications in the `bibliography.bib` file for papers on **cognitive maps**.
The matching papers are listed in Figure \@ref(fig:cognitive-map).
The filter searches for the keyword `cognitive map` in the `title`, `abstract` or `tags` of the bibliography entries.
`r nrow(bib_cognitive_map)` matching papers were found.

```{r, cognitive-map, fig.cap="Table listing bibliography entries containing `cognitive map`"}
DT::datatable(
  data = bib_cognitive_map,
  class = "cell-border stripe", rownames = FALSE, escape = FALSE,
  options = list(
    scrollX = TRUE,
    pageLength = 5,
    lengthMenu = c(5, 10, 15, 20)
  ),
) 
```

## Remapping

```{r}
search_terms = "remapping"
bib_remapping = bib %>%
  .[, ":="(abstract_lower = tolower(abstract), title_lower = tolower(title))] %>%
  # search for search terms in title and abstract
  filter_at(
    .vars = vars(abstract_lower, title_lower, tags),
    .vars_predicate = any_vars(
      stringr::str_detect(
        string = .,
        pattern = paste(search_terms, collapse = "|")))) %>%
  setDT(.) %>%
  .[,c("ID", "title", "author", "year", "journal", "url")]
```

In this section, I filter publications in the `bibliography.bib` file for papers on **remapping** in the hippocampus.
The matching papers are listed in Figure \@ref(fig:remapping).
The filter searches for the keyword `remapping` in the `title`, `abstract` or `tags` of the bibliography entries.
`r nrow(bib_remapping)` matching papers were found.

```{r, remapping, fig.cap="Table listing bibliography entries containing `remapping`"}
DT::datatable(
  data = bib_remapping,
  class = "cell-border stripe", rownames = FALSE, escape = FALSE,
  options = list(
    scrollX = TRUE,
    pageLength = 5,
    lengthMenu = c(5, 10, 15, 20)
  ),
) 
```

## Representations

```{r}
search_terms = "representation"
bib_representation = bib %>%
  .[, ":="(abstract_lower = tolower(abstract), title_lower = tolower(title))] %>%
  # search for search terms in title and abstract
  filter_at(
    .vars = vars(tags),
    .vars_predicate = any_vars(
      stringr::str_detect(
        string = .,
        pattern = paste(search_terms, collapse = "|")))) %>%
  setDT(.) %>%
  .[,c("ID", "title", "author", "year", "journal", "url")]
```

In this section, I filter publications in the `bibliography.bib` file for papers on **representations** and **representation learning**.
This is a bit of challenge, because if you would search for "representation" in paper titles or abstracts, you would end up with a *long* list of papers, since the term is quite ubiquitous in neuroscience, psychology and machine learning.
Therefore, I here resort to manually annotating publications using the `tags` field of the bibliography entries.
The filter shown here searches for the keyword `representation` in the `tags` field of the bibliography entries.
The matching papers are listed in Figure \@ref(fig:representation).
In total, `r nrow(bib_representation)` matching papers were found.

```{r, representation, fig.cap="Table listing bibliography entries containing `representation` in the `tags` field"}
DT::datatable(
  data = bib_representation,
  class = "cell-border stripe", rownames = FALSE, escape = FALSE,
  options = list(
    scrollX = TRUE,
    pageLength = 5,
    lengthMenu = c(5, 10, 15, 20)
  ),
) 
```

## Representational Drift

```{r}
search_terms = "representational drift"
bib_representational_drift = bib %>%
  .[, ":="(abstract_lower = tolower(abstract), title_lower = tolower(title))] %>%
  # search for search terms in title and abstract
  filter_at(
    .vars = vars(abstract_lower, title_lower, tags),
    .vars_predicate = any_vars(
      stringr::str_detect(
        string = .,
        pattern = paste(search_terms, collapse = "|")))) %>%
  setDT(.) %>%
  .[,c("ID", "title", "author", "year", "journal", "url")]
```

In this section, I filter publications in the `bibliography.bib` file for papers on **representational drift**.
The filter shown here searches for the keyword `representation` in the `abstract`, `title` and `tags` fields of the bibliography entries.
The matching papers are listed in Figure \@ref(fig:representational-drift).
In total, `r nrow(bib_representational_drift)` matching papers were found.

```{r, representational-drift, fig.cap="Table listing bibliography entries containing `representational drift` in the `title`, `abstract` and `tags` field"}
DT::datatable(
  data = bib_representational_drift,
  class = "cell-border stripe", rownames = FALSE, escape = FALSE,
  options = list(
    scrollX = TRUE,
    pageLength = 5,
    lengthMenu = c(5, 10, 15, 20)
  ),
) 
```

## Zoo

```{r}
search_terms = "zoo"
bib_zoo = bib %>%
  # search for search terms in title and abstract
  filter_at(
    .vars = vars(tags),
    .vars_predicate = any_vars(
      stringr::str_detect(
        string = .,
        pattern = paste(search_terms, collapse = "|")))) %>%
  setDT(.) %>%
  .[,c("ID", "title", "author", "year", "journal", "url")]
```

Figure \@ref(fig:zoo) shows relevant references for the Zoo project.

```{r, zoo, fig.cap="Table listing bibliography entries containing `zoo` in the `tags` field"}
DT::datatable(
  data = bib_zoo,
  class = "cell-border stripe", rownames = FALSE, escape = FALSE,
  options = list(
    scrollX = TRUE,
    pageLength = 5,
    lengthMenu = c(5, 10, 15, 20)
  ),
) 
```


## Missing PDFs

Figure \@ref(fig:missing-pdfs) lists all publications with missing PDFs.

```{r, missing-pdfs, fig.cap="Table of all publications with a missing PDF."}
bib_pdf = bib %>%
  filter(get("bdsk-file-1") == "") %>%
  filter(journal != "Zenodo") %>%
  setDT(.) %>%
  .[,c("ID", "title", "author", "year", "journal", "url")]
DT::datatable(
  bib_pdf,
  class = "cell-border stripe", rownames = FALSE, escape = FALSE,
  options = list(
    scrollX = TRUE,
    pageLength = 5,
    lengthMenu = c(5, 10, 15, 20)
    ),
  ) 
```

## Missing abstracts

In some of the sections above, I use information from the abstract to filter publications for specific keywords.
It is therefore important that the `abstract` information is complete.
In Figure \@ref(fig:missing-abstract) I filter for all publications with missing `abstract` information to continously update this information.

```{r, missing-abstract, fig.cap="Table of all publications with a missing abstract."}
bib_abstract = bib %>%
  filter(get("abstract") == "") %>%
  filter(journal != "Zenodo") %>%
  setDT(.) %>%
  .[,c("ID", "title", "author", "year", "journal", "url")]
DT::datatable(
  bib_abstract,
  class = "cell-border stripe", rownames = FALSE, escape = FALSE,
  options = list(
    scrollX = TRUE,
    pageLength = 5,
    lengthMenu = c(5, 10, 15, 20)
    ),
  ) 
```

## Publications added per day

```{r}
bib_time = bib %>%
  setorder(date_added_converted) %>%
  .[, by = .(date_added_converted), .(num_added = .N)] %>%
  verify(.[, .(N = .(sum(num_added)))]$N == nrow(bib))
```

Figure \@ref(fig:pub-day) shows the number of publications added to the bibliography per day.
The maximum number of publications added on a single day were `r max(bib_time$num_added)` publications added on `r format(bib_time$date_added_converted[which.max(bib_time$num_added)], "%d %B, %Y")`.

```{r, pub-day, fig.cap="Number of publications added to the bibliography per day (y-axis) over time (in months; x-axis) since creating the bibliography in BibDesk in Sep 2019.", fig.fullwidth=TRUE}
ggplot(data = bib_time, aes(x = date_added_converted, y = num_added)) +
  geom_bar(stat = "identity") +
  ggtitle("Number of publications added to the bibliography per day") +
  ylab("Number of publications added per day") +
  xlab("Time (months)") +
  theme(panel.background = element_blank()) +
  theme(panel.grid.major = element_blank()) +
  theme(panel.grid.minor = element_blank()) +
  theme(axis.line = element_line(colour = "black")) +
  coord_capped_cart(bottom = "both", left = "both") +
  scale_x_date(date_breaks = "1 months", date_labels =  "%b %Y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5))
  #annotate("segment", x = as.Date("2020-11-01 UTC"),
  #         xend = as.Date("2020-12-01 UTC"), y = 45, yend = 45,
  #         colour = "red", arrow = arrow(length = unit(0.30, "cm"))) +
  #geom_label(aes(x = as.Date("2020-10-01 UTC"), y = 45,
  #      label = "hectic review\nwriting starts"),
  #      color = "red", size = 3, fontface = "bold")
```

# Contact

If a reference is missing, please [create a new issue](https://github.com/lnnrtwttkhn/bibliography/issues/new/choose) and use the issue template for missing publications.

If you have any questions about the bibliography, [the repository](https://github.com/lnnrtwttkhn/bibliography/), if you spotted a bug or would like to make a comment, please also [open an issue](https://github.com/lnnrtwttkhn/bibliography/issues/new) first, or otherwise [email Lennart](mailto:wittkuhn@mpib-berlin.mpg.de).

Thanks!

# Behind the scenes

The Python script below (`parser.py`) reads the `bibliography.bib` file and uses (1) `bibtexparser` to read the bibliography contents and (2) `pandas` to transform the content into a `bibliography.csv` file that is read into this notebook.

```{python, echo=TRUE, code=readLines(here::here("code", "parser.py")), eval=FALSE, python.reticulate=FALSE, class.source = 'fold-show'}
```

The `R Markdown` notebook is then rendered on every push to the repo using continuous integration via Travis:

```{bash, echo=TRUE, code=readLines(here::here(".travis.yml")), eval=FALSE, class.source = 'fold-show'}
```

The continuous integration executes a simple `Makefile` that first runs the `parser.py` and then the `bibliography.Rmd` notebook:

```{bash, echo=TRUE, code=readLines(here::here("Makefile")), eval=FALSE, class.source = 'fold-show'}
```

The dependencies of the Python code are listed in `requirements.txt`:

```{bash, echo=TRUE, code=readLines(here::here("requirements.txt")), eval=FALSE, class.source = 'fold-show'}
```

This R Markdown notebook was built using the following computational environment:

```{r, results="hold"}
# set back to old locale
Sys.setlocale("LC_TIME", loc) 
sessionInfo()
```


